---
title: "Reproducible Research: Peer Assesment 1"
author: "Mike Ceballos"
output:
  html_document:
    keep_md: true
---

This report shows how was Peer Assesment 1 executed. It contains an explanation of code required, as well as results obtained during running time.

##Loading and preprocessing the data
It is necessary to set work directory previous to load data.Then data will be loaded in "activity" variable:
```{r}
setwd("C:/Users/204024285/Documents/MisPrimerosProyectos/Reproducible Research")
activity<-read.csv("activity.csv")
```
Next step is to review data structure and get a summary of current information. 
```{r}
str(activity)
summary(activity)
```
At this point, it is able to make the analysis without further changes in current dataset.

## What is mean total number of steps taken per day?
According with assignment, I calculated the total number of steps taken per day, using tapply function as follows:
```{r}
data_hist<-tapply(activity$steps, activity$date, sum)
data_hist
```
Now, an histogram is created to review data distribution.
```{r echo=FALSE, fig.height=4, results='asis'}
hist(data_hist, main= paste("Histogram of Steps per day"))
totalmean <- mean(data_hist, na.rm=TRUE)
```

Finally, the mean total of steps taken per day, without taking in consideration NA values, is: `r totalmean`.

## What is the average daily activity pattern?

In order to solve this question, it is necessary to remove NA values and then calculate information asked by time interval. Finally, a time series plot is presented.
```{r}
data_interval<-na.omit(activity)
dataplot <- tapply(data_interval$steps, data_interval$interval, mean)
plot(as.vector(dataplot), x=names(dataplot), type="l")
index<-which.max(as.vector(dataplot))
time_interval<-names(dataplot[index])
steps<-as.vector(dataplot[index])

```

The 5-minute interval with major steps were at `r time_interval` with `r steps` in average.


## Imputing missing values

```{r, echo=FALSE}
totalNA <- sum(is.na(activity$steps))
```
Now, I will review the amount of missing values on this dataset, evaluating the number of rows with NA values.
The total of NAs values is `r totalNA`.

A simple strategy will be used to cover all missing values, taking in count average of steps per day. Therefore, a new dataset will be created similar to original, but filling missing values with average steps. Finally, a histogram will be created to visualize data.
```{r}
data_fixed <- data_hist
data_fixed[is.na(data_fixed)]<-mean(data_hist,na.rm=TRUE)
hist(data_fixed, main= paste("Histogram of Steps per day"))
```


## Are there differences in activity patterns between weekdays and weekends?

A new factor was created to split dataset in week days and weekends. Then, two plots were created to show data regarding time interval exercise.

```{r}
library(timeDate)
isWeekday(as.Date(names(data_fixed)))
data_interval$weekDay<-isWeekday(as.Date(data_interval$date))
weekDayData<- subset(data_interval, data_interval$weekDay)
weekEndData<- subset(data_interval, !data_interval$weekDay)
dataplot1<-tapply(weekDayData$steps, weekDayData$interval, mean)
dataplot2<-tapply(weekEndData$steps, weekEndData$interval, mean)
plot(as.vector(dataplot1), x=names(dataplot1), type="l")
plot(as.vector(dataplot2), x=names(dataplot2), type="l")
```


For any comments, please don't hesitate to contact me (Miguel Ceballos).
